---
title: "Central Limit Theory and the Exponential Distribution"
author: "Andrew Crisp"
date: "November 13, 2015"
output: pdf_document
---

# Overview
Through a simulation, the Central Limit Theory will be demonstrated.  The mean and variance of the exponential distribution will be simulated and compared to calculated values.

# Simulations
During the simulation, lambda will be set to a controlled 0.2.  

```{r}
lambda <- 0.2
nosims <- 1000
means <- NULL
rexpValues <- NULL
n <- 40
rexpValues <- rexp(n = nosims, rate = lambda)
#for (i in 1:nosims) rexpValues <- c(rexpValues,rexp(n = n, rate = lambda))
```

# Sample mean versus Theoretical mean
The mean of an exponential distribution is 1/lambda.  This theoretical mean can be calculated easily and compared with the actual collected sample mean.  

```{r}
calcMean <- 1/lambda
n40CLTMean <- mean(sample(rexpValues, n, replace = FALSE))
n100CLTMean <- mean(sample(rexpValues, 100, replace = FALSE))
simMean <- mean(rexpValues)
```

The calculated mean is `r calcMean`.

The mean according to Central Limit Theory with a sample size of `r n` should be near `r n40CLTMean`.  With a sample size of 100, the mean should be near `r n100CLTMean`.  The larger sample size will help the mean approach the actual mean of the simulated values.

The actual mean of the simulated values is `r simMean`.

# Sample variance versus Theoretical variance
Similar to the mean, the standard deviation of an exponential distribution is 1/lambda.  This theoretical mean can be calculated easily and compared with the actual collected sample mean.  


# Distribution
