---
title: "Central Limit Theory and the Exponential Distribution"
author: "Andrew Crisp"
date: "November 13, 2015"
output: pdf_document
---

# Overview
Through a simulation, the Central Limit Theory will be demonstrated.  The mean and variance of the exponential distribution will be simulated and compared to calculated values.

# Simulations
During the simulation, lambda will be set to a controlled 0.2.  1000 passes will be made, collecting 40 random exponential distribution samples.  After collecting the sample values, the mean of each group will be collected and indexed for easy inspection.

```{r}
lambda <- 0.2
nosims <- 1000
n <- 40
rexpValues <- NULL
means <- NULL
for (i in 1:nosims) rexpValues <- rbind(rexpValues,rexp(n = n, rate = lambda))
for (i in 1:nosims) means <- rbind(means, mean(rexpValues[i,]))
means <- as.numeric(means)
```

# Sample mean versus Theoretical mean
The mean of an exponential distribution is 1/lambda.  This theoretical mean can be calculated easily and compared with the actual collected sample mean.  

```{r}
calcMean <- 1/lambda
simMean <- mean(rexpValues)
CLTMean <- mean(means)
```

The calculated mean is `r calcMean`.

The Central Limit Theory says that the mean of the a sample should approximate the mean of the population.  The mean of all 1000 means simulated is `r CLTMean`.

The actual mean of all the simulated values is `r simMean`.

```{r}
require(ggplot2)
ggplot() +
  aes(means) + 
  geom_histogram(aes(y = ..density..), binwidth=.25, alpha=.2) + 
  geom_vline(xintercept=mean(means)) +
  stat_function(geom="line", fun=dnorm, arg=list(mean=mean(means),sd=sd(means)))
```

# Sample variance versus Theoretical variance
Similar to the mean, the standard deviation of an exponential distribution is 1/lambda.  The variance is the square of the standard deviation.  However, the variance of the samples should approach that of a normal distribution.  

```{r}
calcVariance <- (1/lambda)^2
n <- 40
n40CollectionOfVariances <- NULL
for (i in 1:nosims) n40CollectionOfVariances <- rbind(n40CollectionOfVariances, var(rexpValues[i,]))
simVariance <-mean(n40CollectionOfVariances)
```

The calculated (theoretical) variance is `r calcVariance`.

The average variance of the simulations is `r simVariance`.

```{r}
ggplot() +
  aes(n40CollectionOfVariances) + 
  geom_histogram(aes(y = ..density..), alpha=.8, binwidth=1) + 
  geom_vline(xintercept=mean(n40CollectionOfVariances)) +
  stat_function(geom="line", fun=dnorm, arg=list(mean=mean(n40CollectionOfVariances),sd=sd(n40CollectionOfVariances))) + 
  geom_histogram(aes(rexpValues[,1],y = ..density..), alpha=.2, binwidth=1) + 
  stat_function(geom="line", fun=dexp,arg=list(rate=lambda))
```

# Distribution

