---
title: "Tooth Growth Analysis"
author: "Andrew Crisp"
date: "November 18, 2015"
output: html_document
---

# Load the data
```{r message=FALSE}
library(datasets)
library(ggplot2)
library(dplyr)
```

The ToothGrowth dataset is packaged with the R download.  This dataset shows the effect of vitamin C on tooth growth in guinea pigs.  The vitamin C is delivered either through orange juice or ascorbic acid.  

# Initial exploration

Before evaluating the data, an initial familiarization of the information is necessary.

```{r}
str(ToothGrowth)
summary(ToothGrowth)
```
The structure shows three variables.  Further, the dose looks like it has only a few discrete values.  The assumption can be made that this is a factor.  After converting to a factor, a cursory glance suggests that orange juice encourages more growth than ascorbic acid, at least in the smaller doses.

```{r}
ToothGrowth$dose <- factor(ToothGrowth$dose)
ToothGrowth %>% tbl_df %>% group_by(dose,supp) %>% summarise_each(funs(mean))
ggplot(ToothGrowth, aes(x=rev(supp), y=len, color=dose)) + geom_boxplot()

```

# Growth by treatment

## Dose size .5
```{r}
t.test(filter(ToothGrowth,supp=='OJ',dose==.5)$len,filter(ToothGrowth,supp=='VC',dose==.5)$len)
```

At dose size 0.5, the .95 confidence interval shows the real difference in the mean about 1.719 - 8.871.  Accordingly, it can be said with 95% confidence that the mean tooth growth is larger when dosed at 0.5 level with acerbic acid.

## Dose size 1
```{r}
t.test(filter(ToothGrowth,supp=='OJ',dose==1)$len,filter(ToothGrowth,supp=='VC',dose==1)$len)
```

At dose size 1.0, the .95 confidence interval shows the real difference in the mean about 2.802 - 9.058.  Accordingly, it can be said with 95% confidence that the mean tooth growth is larger when dosed at 1.0 level with acerbic acid.

## Dose size 2
```{r}
t.test(filter(ToothGrowth,supp=='OJ',dose==2)$len,filter(ToothGrowth,supp=='VC',dose==2)$len)
```

At dose size 2.0, the .95 confidence interval shows the real difference in the mean about -3.798 - 3.638.  Accordingly, it cannot be said that there is a difference in tooth growth between the two delivery methods at the 2.0 dose level.