---
title: "Tooth Growth Analysis"
author: "Andrew Crisp"
date: "November 18, 2015"
output: html_document
---

# Load the data
```{r message=FALSE}
library(datasets)
library(ggplot2)
library(dplyr)
```

The ToothGrowth dataset is packaged with the R download.  This dataset shows the effect of vitamin C on tooth growth in guinea pigs.  The vitamin C is delivered either through orange juice or ascorbic acid.  

# Initial exploration

Before evaluating the data, an initial familiarization of the information is necessary.

```{r}
str(ToothGrowth)
summary(ToothGrowth)
```
The structure shows three variables.  Further, the dose looks like it has only a few discrete values.  The assumption can be made that this is a factor.  After converting to a factor, a cursory glance suggests that orange juice encourages more growth than ascorbic acid, at least in the smaller doses.

```{r}
ToothGrowth$dose <- factor(ToothGrowth$dose)
ToothGrowth %>% tbl_df %>% group_by(supp,dose) %>% summarise_each(funs(mean))
ggplot(ToothGrowth, aes(x=supp, y=len, group=dose, color=dose)) + geom_line() + geom_point()

```

# Growth by treatment
